---
title: "Homework 6 Report"
author: "Isabelle Meredith"
date: "`r Sys.Date()`"

knit: (function(inputFile, encoding) {
      out_dir <- "../Reports";
      rmarkdown::render(inputFile,
                        output_format="pdf_document",
                        encoding=encoding,
                        output_dir=out_dir)}) #file.path(dirname(inputFile), out_dir)
---


```{r, include = FALSE}
# HW 6 Script

# Author: Isabelle Meredith
# Date: 10/3/2024
# Description: Script to evaluate and plot the impact of nearby lead smelter on children's IQ

library(knitr)
library(ggplot2)
library(dplyr)
```

```{r}
dfLeadRaw <- read.csv("../DataRaw/lead-iq-01.csv")

head(dfLeadRaw)

# Data processing takes place here. Surely in the future I won't need to do anything with this
write.csv(dfLeadRaw, "../DataProcessed/lead-iq-cleaned.csv")

# Read in the processed form of the data instead
dfLead <- read.csv("../DataProcessed/lead-iq-cleaned.csv")
  
```

```{r}
ggplot(dfLead, aes(IQ, fill = Smelter)) + 
   geom_histogram(alpha = 0.6, position = 'identity', binwidth = 5) + 
   xlab("Child IQ") +
   guides(fill = guide_legend(title = "Distance to Smelter")) +
   theme(legend.position = "top")
```
Histogram of IQs comparing children who were near to the smelter with children who were far from the smelter. 

```{r}
ggplot(dfLead, aes(x=IQ, y = Smelter, fill = Smelter)) + 
   geom_boxplot(position = 'identity') +
   ylab("Distance to Smelter") +
   guides(fill="none")
```
Boxplot of IQs comparing children who were near to the smelter with children who were far from the smelter. 

```{r, include = FALSE}
meanNear <- mean(dfLead[dfLead$Smelter == "Near", "IQ"])
meanFar <- mean(dfLead[dfLead$Smelter == "Far", "IQ"])
```

```{r}
dfLeadSummary <- dfLead %>% 
  group_by(Smelter) %>%
  rename("Smelter Distance" = Smelter) %>%
  summarise("Mean IQ" = mean(IQ), "IQ SD" = sd(IQ), "Group Count" = n())
kable(dfLeadSummary)
```
The table contains the values for the group means, standard deviations, and counts for the group near and far from the smelter.

The mean IQ for close to the smelter is `r meanNear` and the mean IQ far from the smelter is `r meanFar`. 